---
title: "MultipleImputation_SpecificGravity"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Install and load required packages
```{r}
#install.packages("mice")
#install.packages("lubridate")
#install.packages("lattice")
#install.packages("gamm4")
#install.packages("Hmisc")
#install.packages("digest")
library(haven)
library(data.table)
library(ggplot2)
library(nlme)
library(mgcv)
library(gamm4)
library(lattice)
library(lubridate)
library(digest)
devtools::install_github("ja-thomas/autoxgboost") 
library(autoxgboost)
library(xgboost)
#install.packages("DiagrammeR")
library(DiagrammeR)

#install.packages(c("NMF", "RANN", "fastICA", "RcppRoll", "Matrix", "ipred", "recipes"))
#library(c(NMF, RANN, fastICA, RcppRoll, Matrix, ipred, recipes))
#install.packages("prodlim")

#install.packages("caret")
library(caret)

#install.packages("robustbase")
library(robustbase)
#install.packages("randomForest")
library(randomForest)


```

## Set path and import data
```{r}
path = '/Users/dewate01/Desktop/PROGRESS_SpecificGravity_MultipleImputation/PROGRESS_SpecificGravity_MultipleImputation'

#dataset downloaded from Box folder, "Specific Gravity Missing Data Imputation"
#In it, Erik merged previous variables with additional variables that I sent
SG_data <- read.csv("ERIK_NIA_SG_imp_nov7_2018_updated.csv")
```

##compute season of collection date to control for seasonal variation in data collection. 
```{r}
SG_data$date = mdy(SG_data$fecha_control)
SG_data$days = yday(SG_data$date) 

```

## create an average variable that represents the geometric mean of 5 phthalates that are weakly correlated with each other (as recommendended by Dr. Alan Just). Associations between this average phthalates variable and specific gravity will later be explored
```{r}
#create natural log of phtalates of interest
SG_data$mEP2_Conc_ngml_l <- log(SG_data$mEP2_Conc_ngml)
SG_data$mBP_Conc_ngml_l <- log(SG_data$mBP_Conc_ngml)
SG_data$miBP_Conc_ngml_l <- log(SG_data$miBP_Conc_ngml)
SG_data$mBzP2_Conc_ngml_l <- log(SG_data$mBzP2_Conc_ngml)
SG_data$mECPP_Conc_ngml_l <- log(SG_data$mECPP_Conc_ngml)

#create exponential function then create variable which is geometric mean
e <- exp(1) 
SG_data$phthalates <-e^(rowMeans(SG_data[, c("mEP2_Conc_ngml_l", "mBP_Conc_ngml_l", "miBP_Conc_ngml_l", 
                                               "mBzP2_Conc_ngml_l", "mECPP_Conc_ngml_l")]))
```


##Check distribution of specific gravity. Check for normal distribution and outliers
```{r echo= FALSE}
#plot a histogram to check distribution
ggplot(data = SG_data, aes(x = SG, fill = "Red")) +geom_histogram()

#compute mean and SD and check whether there are SG values >3 SD's above/below mean
mean_SG = mean(SG_data$SG, na.rm =T)
sd_SG = sd(SG_data$SG, na.rm =T)
SD3_below = mean_SG - (3*sd_SG)
SD3_above = mean_SG + (3*sd_SG)
minSG = min(SG_data$SG, na.rm=T)
maxSG = max(SG_data$SG, na.rm=T)
SG_data$SG_outlier [SG_data$SG> SD3_above] <-1
SG_data$SG_outlier [SG_data$SG< SD3_above] <-0

```

#run xgboost using training dataset with subjects that have SG that we artificially applied missing values 
```{r, echo= FALSE}

run.autoxgboost.dart <- function(traindt, target, ...){
  require(autoxgboost)
  par_dart <- makeParamSet(
    makeNumericParam("eta", lower = 0.01, upper = 0.2),
    makeNumericParam("gamma", lower = -7, upper = 6, trafo = function(x) 2^x),
    makeIntegerParam("max_depth", lower = 3, upper = 20),
    makeNumericParam("colsample_bytree", lower = 0.5, upper = 1),
    makeNumericParam("colsample_bylevel", lower = 0.5, upper = 1),
    makeNumericParam("lambda", lower = -10, upper = 10, trafo = function(x) 2^x),
    makeNumericParam("alpha", lower = -10, upper = 10, trafo = function(x) 2^x),
    makeNumericParam("subsample", lower = 0.5, upper = 1),
    makeDiscreteParam("booster", values = c("dart")),
    makeDiscreteParam("sample_type", values = c("uniform", "weighted"), 
                      requires = quote(booster == "dart")),
    makeDiscreteParam("normalize_type", values = c("tree", "forest"), 
                      requires = quote(booster == "dart")),
    makeNumericParam("rate_drop", lower = 0, upper = 1, requires = quote(booster == "dart")),
    makeNumericParam("skip_drop", lower = 0, upper = 1, requires = quote(booster == "dart")),
    makeLogicalParam("one_drop", requires = quote(booster == "dart"))
  )
  
  reg_task <- makeRegrTask(data = as.data.frame(traindt), target = target)
  MBOctrl <- makeMBOControl()
  ctrl <- setMBOControlTermination(control = MBOctrl, iters = 160)
  
  set.seed(1234)
  reg_auto_dart <- autoxgboost(reg_task, par.set = par_dart,
                               control = ctrl
                               # nthread = xgb_threads
  )
  return(reg_auto_dart)
}

SG_data<-setDT(SG_data)
SG_nomiss <- SG_data[!is.na(SG) & !is.na(phthalates)]


# create 5 folds

 
n_folds <- 5
SG_nomiss$index <- 1:nrow(SG_nomiss)
Folds <- caret::createFolds(SG_nomiss$index, k = n_folds)
str(Folds)

# store output
rmse_cv <- rep(NA, n_folds)
dart_parameters <- dart_parameters2 <-  list()
 install.packages("paralell")
 library(parallel)
parallel::mclapply(1:5,function(i){
  tmp = SG_nomiss[, .(gest_age_weeks, SG, days,mothers_age, mother_bmi, phthalates)]
  train_XY <- tmp[-Folds[[i]] ]
  test_XY <-  tmp[ Folds[[i]] ]
  # this step takes a while.
  reg_auto_dart <- run.autoxgboost.dart(traindt = train_XY, target = 'SG')
  # predict to test dataset
  xgb_pred <- predict(reg_auto_dart, newdata=test_XY)
  SG_nomiss$pred[Folds[[i]] ] <<- xgb_pred$data$response
  rmse_cv[i] <<- sqrt(mean((xgb_pred$data$truth - xgb_pred$data$response)^2))  
  
  # I stored two versions of parameters
  dart_parameters[[i]] <<- mlr::getHyperPars(reg_auto_dart$final.learner)
  dart_parameters2[[i]] <<- unlist(dart_parameters[[i]])
}, mc.cores=1)




# total rmse is:
rmse_total <- sqrt(mean(rmse_cv^2))
rmse_total

#create different subsets for each fold, in order to compute R2 per fold

SG_fold1 = subset(SG_nomiss, index %in% Folds$Fold1)
SG_fold2 = subset(SG_nomiss, index %in% Folds$Fold2)
SG_fold3 = subset(SG_nomiss, index %in% Folds$Fold3)
SG_fold4 = subset(SG_nomiss, index %in% Folds$Fold4)
SG_fold5 = subset(SG_nomiss, index %in% Folds$Fold5)

R2_fold1 = (cor(SG_fold1$SG, SG_fold1$pred))^2
R2_fold2 = (cor(SG_fold2$SG, SG_fold2$pred))^2
R2_fold3 = (cor(SG_fold3$SG, SG_fold3$pred))^2
R2_fold4 = (cor(SG_fold4$SG, SG_fold4$pred))^2
R2_fold5 = (cor(SG_fold5$SG, SG_fold5$pred))^2

# R2 (how much has been explained): 

R2 = (1- rmse_total / sqrt(mean((SG_nomiss$SG - mean(SG_nomiss$SG))^2)))
R2

#R2 of correlation between observed and predicted SG values
R2_pred = (cor(SG_nomiss$SG, SG_nomiss$pred))^2
R2_pred


summary(SG_nomiss$pred)
```
##compute density plot (xgboost)
```{r, echo= FALSE}
#create subsets for observed and predicted SG values

SG_nomiss<-setDT(SG_nomiss)

observed_xg<- SG_nomiss[,.(SG)]
predicted_xg<- SG_nomiss[,.(pred)]


#create indicator variables for each predicted and observed SG #dataset
observed_xg[!is.na(SG), SG_cat:= 1]
predicted_xg[!is.na(pred), SG_cat:= 2]
setnames(predicted_xg, "pred", "SG")


#combine observed and predicted values so that they both
#are in the same variable
SG_imp_xg <- rbind(observed_xg, predicted_xg)

#give 'factor' class to indicator variable
SG_imp_xg$SG_cat <-as.factor(SG_imp$SG_cat)
SG_imp_xg[SG_cat== 1, SG_group:= "Observed Specific Gravity"]
SG_imp_xg[SG_cat== 2, SG_group:= "Imputed Specific Gravity"]
SG_imp_xg$SG_group <- as.factor(SG_imp$SG_group)


#density plot for observed and predicted SG
ggplot(SG_imp_xg, aes(x = SG, color = SG_group)) + geom_density(alpha = 0.5) +
  geom_hline(yintercept=0, colour="white", size=1) +
  ylab("Density")+
  xlab("Specific Gravity") +
  labs(color = "Specific Gravity Type") +
  theme(legend.title = element_text(size = 10, face = "bold")) + ggtitle("Density Plot of Observed and Imputed Specific Gravity (xgboost)") +
  theme(plot.title = element_text(size = 12, face = "bold")) +
  theme_bw()

```
##Use autoxgboost to predict missing values (change training set to subjects with measured SG, and testing set to those with missing values)
```{r}
run.autoxgboost.dart <- function(traindt, target, ...){
  require(autoxgboost)
  par_dart <- makeParamSet(
    makeNumericParam("eta", lower = 0.01, upper = 0.2),
    makeNumericParam("gamma", lower = -7, upper = 6, trafo = function(x) 2^x),
    makeIntegerParam("max_depth", lower = 3, upper = 20),
    makeNumericParam("colsample_bytree", lower = 0.5, upper = 1),
    makeNumericParam("colsample_bylevel", lower = 0.5, upper = 1),
    makeNumericParam("lambda", lower = -10, upper = 10, trafo = function(x) 2^x),
    makeNumericParam("alpha", lower = -10, upper = 10, trafo = function(x) 2^x),
    makeNumericParam("subsample", lower = 0.5, upper = 1),
    makeDiscreteParam("booster", values = c("dart")),
    makeDiscreteParam("sample_type", values = c("uniform", "weighted"), 
                      requires = quote(booster == "dart")),
    makeDiscreteParam("normalize_type", values = c("tree", "forest"), 
                      requires = quote(booster == "dart")),
    makeNumericParam("rate_drop", lower = 0, upper = 1, requires = quote(booster == "dart")),
    makeNumericParam("skip_drop", lower = 0, upper = 1, requires = quote(booster == "dart")),
    makeLogicalParam("one_drop", requires = quote(booster == "dart"))
  )
  
  reg_task <- makeRegrTask(data = as.data.frame(traindt), target = target)
  MBOctrl <- makeMBOControl()
  ctrl <- setMBOControlTermination(control = MBOctrl, iters = 160)
  
  set.seed(1234)
  reg_auto_dart <- autoxgboost(reg_task, par.set = par_dart,
                               control = ctrl
                               # nthread = xgb_threads
  )
  return(reg_auto_dart)
}

#subset main dataset to only variables of interest
SG_data<-setDT(SG_data)
SG_data<- SG_data[, .(folio, gest_age_weeks, SG, mothers_age, mother_bmi, days, phthalates)]

#subset only to subjects with complete SG and phtalates data
SG_nomiss <- SG_data[!is.na(SG) & !is.na(phthalates)]

#subset only to subjects with missing SG and complete phtalates data
SG_miss <- SG_data[is.na(SG) & !is.na(phthalates)]

#### CREATE FOLDS ####

# create 5 folds for dataset of subjects with complete SG
n_folds <- 5
SG_nomiss$index <- 1:nrow(SG_nomiss)
Folds <- caret::createFolds(SG_nomiss$index, k = n_folds)
str(Folds)

# create 5 folds for dataset of subjects with no SG
SG_miss$index <- 1:nrow(SG_miss)
Folds2 <- caret::createFolds(SG_miss$index, k = n_folds)
str(Folds2)

dart_parameters <- dart_parameters2 <-  list()
install.packages("paralell")
library(parallel)
parallel::mclapply(1:5,function(i){
  tmp = SG_nomiss[, .(gest_age_weeks, SG, days, mothers_age, mother_bmi, phthalates)]
  tmp2 = SG_miss[, .(gest_age_weeks, SG, days, mothers_age, mother_bmi, phthalates)]
  train_XY <- tmp[ Folds[[i]] ]
  test_XY <-  tmp2[ Folds2[[i]] ]
  # this step takes a while.
  reg_auto_dart <- run.autoxgboost.dart(traindt = train_XY, target = 'SG')
  # predict to test dataset
  xgb_pred <- predict(reg_auto_dart, newdata=test_XY)
  SG_miss$pred [Folds2[[i]] ] <<- (xgb_pred$data$response)
  rmse_cv <<- sqrt(mean((xgb_pred$data$truth - xgb_pred$data$response)^2))  
  
  # I stored two versions of parameters
  dart_parameters[[i]] <<- mlr::getHyperPars(reg_auto_dart$final.learner)
  dart_parameters2[[i]] <<- unlist(dart_parameters[[i]])
}, mc.cores=1)#mc.cores should be set to 1 on Windows and 5 on Mac


#create indicator variable of  imputed and non imputed
SG_nomiss[, imp_indic := 0]
SG_miss[, imp_indic := 1]

#create indicator variables that specifies what 0/1 categories #represent
SG_nomiss[imp_indic == 0, imp_group := "Original values"]
SG_miss[imp_indic == 1, imp_group := "Imputed values"]

#drop SG that has all missing
SG_miss[, SG := NULL]

#give predicted values the column name of SG
#so that it will be inserted into same column as observed SG
#when you do rbind
setnames(SG_miss, "pred", "SG")


#combine datasets with missing and nonmissing SG
final <- rbind(SG_nomiss, SG_miss)
final <- as.data.table(final)


#Export final dataset: final
write.csv(final, file = "J:/PM/Element/ANALYSIS/Specific Gravity Missing Data Imputation/SG Imputation/PROGRESS_SpecificGravity_MultipleImputation/final_SG_imputation_nov18_2018.csv", row.names = F)


#Scatterplot with seperate smoothing lines for observed and predicted SG
ggplot(final, aes(x=phthalates, y=SG, color=imp_group)) +
  geom_point(alpha = 0.15, position=position_jitter(width=1,height=.1)) +
  stat_smooth(method="lm", se = FALSE) +
  theme_bw() +
  labs(color = "Specific Gravity Type") +
  theme(legend.title = element_text(size = 10, face = "bold")) +
  ylab("Specific Gravity") +
  xlab("Phthalates") +
  ggtitle("Scatterplot of Observed and Predicted Specific Gravity") +
  theme(plot.title = element_text(size = 12, face = "bold"))    



```


```{r, echo= FALSE}

