---
title: "MultipleImputation_SpecificGravity"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Install and load required packages
```{r}
install.packages("mice")
install.packages("lubridate")
install.packages("lattice")
install.packages("gamm4")
library(haven)
library(mice)
library(data.table)
library(ggplot2)
library(nlme)
library(mgcv)
library(gamm4)
library(lattice)
library(lubridate)
```

## Set path and import data
```{r}
path = '/Users/dewate01/Desktop/PROGRESS_SpecificGravity_MultipleImputation/PROGRESS_SpecificGravity_MultipleImputation'
setwd(path)
SG_data = read.csv('ERIK_NIA_SG_imp_nov7_2018_updated.csv')
```

##compute season of collection date to control for seasonal variation in data collection. First convert the date into a single number, and then compute the sine and cosine of time as indexes of season
```{r}
SG_data$date = mdy(SG_data$fecha_control)
SG_data$days = yday(date) 
SG_data$time.sin = sinpi(2 * (SG_data$days - 1)/(365 + lubridate::leap_year(SG_data$date) - 1))
SG_data$time.cos = cospi(2 * (SG_data$days - 1)/(365 + lubridate::leap_year(SG_data$date)- 1))

##plot sine and cosine to check distribution
plot_sin = ggplot(SG_data, aes(days, time.sin)) + geom_point() 
plot_sin
plot_cos = ggplot(SG_data, aes(days, time.cos)) + geom_point() 
plot_cos
```

## create an average variable that represents the geometric mean of 5 phthalates that are weakly correlated with each other (as recommendended by Dr. Alan Just). Associations between this average phthalates variable and specific gravity will later be explored
```{r}
#create natural log of phtalates of interest
SG_data$mEP2_Conc_ngml_l <- log(SG_data$mEP2_Conc_ngml)
SG_data$mBP_Conc_ngml_l <- log(SG_data$mBP_Conc_ngml)
SG_data$miBP_Conc_ngml_l <- log(SG_data$miBP_Conc_ngml)
SG_data$mBzP2_Conc_ngml_l <- log(SG_data$mBzP2_Conc_ngml)
SG_data$mECPP_Conc_ngml_l <- log(SG_data$mECPP_Conc_ngml)

#create exponential function then create variable which is geometric mean
e <- exp(1) 
SG_data$phthalates <-e^(rowMeans(SG_data[, c("mEP2_Conc_ngml_l", "mBP_Conc_ngml_l", "miBP_Conc_ngml_l", 
                                               "mBzP2_Conc_ngml_l", "mECPP_Conc_ngml_l")]))
```

##linear regression analyses, regressing covariates and phthalates (geometric mean of 5 phthalates) on specific gravity. These analyses are performed to check whether phthalates and covariates are linearly associated with specific gravity. Covariates were suggested by a clinician (phthalates, maternal age, maternal BMI, season, gestational age), selected based on a literature review on prenatal phthalate exposure (parity (v69), secondhadn smoking, SES, prenatal depression and anxiety, maternal education) and biomarkers of infection (selected by Nia & Erik, given that infection affects urine density).  
```{r}

#multiple regression, excluding variables that have >25% missing data (v69, epds_pregn, sum_pregn_anx)
lm_PhthalatesCovariates = lm(SG_data$SG ~ SG_data$phthalates + SG_data$mothers_age + SG_data$mother_bmi + SG_data$time.sin + SG_data$cbc_basoM + SG_data$smoke_inside + SG_data$ferritina_M + SG_data$cbc_adeM + SG_data$cbc_eosiM + SG_data$cbc_hematM + SG_data$cbc_hemogM + SG_data$cbc_linfoM + SG_data$cbc_monoM + SG_data$cbc_volcorM + SG_data$SES_3cat2T + SG_data$educ_3cat2T + SG_data$gest_age_weeks + SG_data$time.cos)
summary(lm_PhthalatesCovariates)

#3 covariates are associated with SG in linear regression analyses(p<0.1): phtahalates, times.sin and cbc_basoM
```
###  Conduct spline regression to see if nonlinear model is appropriate 

```{r}
#I'm using the thin plate regression spline which is the default as it is most #popularly used spline.  Gestational age, secondhand smoking, SES and maternal education were excluded from this model, because I got the error message: A term has fewer unique covariate combinations than specified maximum degrees of freedom.

non_linear <- gam(SG ~ s(phthalates) + s(mothers_age) + s(mother_bmi)  + s(time.cos) + s(time.sin) + s(cbc_basoM) + s(ferritina_M) + s(cbc_adeM) + s(cbc_eosiM) + s(cbc_hematM) + s(cbc_hemogM) +s(cbc_linfoM) + s(cbc_monoM) +s(cbc_volcorM), data = SG_data)
summary(non_linear)

#variables that are better modeled by a non-linear term (i.e., edf much larger than 1) are phthalates, maternal BMI, ferritin, hemoglobin, lymphocytes, and monocytes

```


##Check distribution of specific gravity. Check for normal distribution and outliers
```{r}
#plot a histogram to check distribution
ggplot(data = SG_data, aes(x = SG, fill = "Red")) +geom_histogram()

#compute mean and SD and check whether there are SG values >3 SD's above/below mean
mean_SG = mean(SG_data$SG, na.rm =T)
sd_SG = sd(SG_data$SG, na.rm =T)
SD3_below = mean_SG - (3*sd_SG)
SD3_above = mean_SG + (3*sd_SG)
minSG = min(SG_data$SG, na.rm=T)
maxSG = max(SG_data$SG, na.rm=T)
SG_data$SG_outlier [SG_data$SG> SD3_above] <-1
SG_data$SG_outlier [SG_data$SG< SD3_above] <-0
#remove outliers from dataset (one subject, folio 380, with biologically impossible value)
SG_data = SG_data[is.na(SG_data$SG) | SG_data$SG < SD3_above,]

```

##Use mice package to perform multiple imputation analyses (10 iterations)
```{r}
#call the mice package
init = mice(SG_data, maxit=0) 
meth = init$method
predM = init$predictorMatrix

#indicate the variables that will be removed as predictors
predM[, c("SG", "folio")]=0


#specify the predictors (norm = continuous variables)
meth[c("phthalates")]="norm"
meth[c("mothers_age")]="norm"
meth[c("mother_bmi")]="norm"
meth[c("gest_age_weeks")]="norm"
meth[c("time.cos")]="norm"
meth[c("time.sin")]="norm"

#perform mi analyses with pmm method
set.seed(103)
imputed_pmm = mice(SG_data, method='pmm', predictorMatrix=predM, m=10)
imputed_pmm <- complete(imputed_pmm)

#Check for missing data in the imputed dataset.
sapply(imputed_pmm, function(x) sum(is.na(x)))

#convert to data table
class(imputed_pmm)
imputed_pmm<-as.data.table(imputed_pmm)


#Density Plot of Original and Imputed Specific Gravity
plot(density(imputed_pmm$SG), col="blue", 
     xlab="Specific Gravity", #Change the x-axis label
     ylab="Density", #y-axis label
     main="Density Plot of Original and Imputed Specific Gravity")#Main title
lines(density(SG_sub$SG, na.rm=TRUE), col="red")
legend(1.04, 50, legend=c("Original SG", "Imputed SG"),
       col=c("red", "blue"), lwd=1:1, cex=0.8,
       title="SG Type", text.font=4, bg='white')


```

