---
title: "Nia_McRae_SG_Impute_oct10_2018"
author: "Nia McRae"
date: "October 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Nia's R Markdown for Multiple Imputation

DATE: October 16, 2018

Nia McRae and Erik de Water created this R markdown for conducting multiple imputation for 100 missing Specific Gravity data

They split up the work. Below is the code for Nia's steps. Erik's steps are in a seperate R markdown document

Erik and Nia are using github to follow each other's work and will eventually merge their R markdowns into one complete document in an effort to produce reproducible code for multiple imputation. 

#First, install and load require packages

```{r variable cleaning and transformations}
#install and load require packages
#install.packages("mice")
#install.packages("lubridate")
#install.packages("lattice")
library(haven)
library(mice)
library(data.table)
library(ggplot2)
library(nlme)
library(mgcv)
library(gamm4)
library(lattice)
library(lubridate)
#lattice package interfaces very well with the “mids” object #type that the mice function returns

```
## Load data and subset to variables you need for the analysis

```{r load data, echo=FALSE}

#load data
SG_data <- read_sas("J:/PM/Element/DATABASES-Mexico/DATASETS SENT/erik_nia_sg_imp_oct10_2018.sas7bdat")

SG_data <-as.data.table(SG_data)

#subset to variables needed for model
SG_sub <- SG_data[,.(folio, gest_age_weeks, etapa, mothers_age, fecha_control, mother_bmi, SG,
mEP2_Conc_ngml, mBP_Conc_ngml, miBP_Conc_ngml, mBzP2_Conc_ngml, mECPP_Conc_ngml)]

SG_sub <- SG_sub[etapa =="2T"]

#see that all phtalates and SG are numeric 
class(SG_sub$mEP2_Conc_ngml)
class(SG_sub$mBP_Conc_ngml)
class(SG_sub$miBP_Conc_ngml)
class(SG_sub$mBzP2_Conc_ngml)
class(SG_sub$mECPP_Conc_ngml)
class(SG_sub$SG)

```
## Compute Geometric Mean

```{r geometric mean, echo=FALSE}

#create natural log of phtalates of interest
SG_sub[, mEP2_Conc_ngml_l := log(mEP2_Conc_ngml)]
SG_sub[, mBP_Conc_ngml_l := log(mBP_Conc_ngml)]
SG_sub[, miBP_Conc_ngml_l := log(miBP_Conc_ngml)]
SG_sub[, mBzP2_Conc_ngml_l := log(mBzP2_Conc_ngml)]
SG_sub[, mECPP_Conc_ngml_l := log(mECPP_Conc_ngml)]

#create exponential function then create variable to represent the geometric #mean of all five phtalates
e <- exp(1) 
SG_sub[, phthalates := e^(rowMeans(SG_sub[, c("mEP2_Conc_ngml_l", "mBP_Conc_ngml_l", "miBP_Conc_ngml_l", 
                                               "mBzP2_Conc_ngml_l", "mECPP_Conc_ngml_l")]))]

```
#create season of data collection variable by converting date of visit into days #of the year and then converting it to sine/cosine

```{r season variable, echo=FALSE}

SG_sub[, date_visit := ymd(fecha_control)]
SG_sub[, days := yday(date_visit) ]

#have maximum be either 365 days or include leap year if there is leap year
SG_sub[, time.sin := sinpi(2 * (days - 1)/(365 + lubridate::leap_year(date_visit)(days, na.rm = T) - 1))]

SG_sub[, time.cos := cospi(2 * (days - 1)/(365 + lubridate::leap_year(date_visit)(days, na.rm = T) - 1))]

#review descriptive stats of specific gravity variable
summary(SG_sub$SG)
sd(SG_sub$SG, na.rm=TRUE)

SG_sub[folio == 380]


#review scatterplots and see if there is any outliers
#It looks like there's one possible outlier which is value of 1.057000
#this is more than 3 SD from the mean
#the mean and median are the same: 1.016 
#and the SD is really small; 0.00660113
ggplot(SG_sub, aes(x=phthalates, y=SG)) + geom_point()
ggplot(SG_sub, aes(x=mothers_age, y=SG)) + geom_point()
ggplot(SG_sub, aes(x=mother_bmi, y=SG)) + geom_point()
ggplot(SG_sub, aes(x=gest_age_weeks, y=SG)) + geom_point()
ggplot(SG_sub, aes(x=time.sin, y=SG)) + geom_point()
ggplot(SG_sub, aes(x=time.cos, y=SG)) + geom_point()

```
#lets find the folio who is the outlier and see if she is outlier in other #places as well (i.e. attendance list and comments)

```{r outlier, echo=FALSE}


#sort in ascending order to see folio with highest SG at very end
SG_sub_ascend <- SG_sub[order(SG),]

#here is the folio with the outlier
outlier_folio <- SG_sub[folio == 380 & etapa == '3T']
#Allan said set this to missing 

#we are only focusing on 2T



```
## Conduct spline regression to see if nonlinear model is appropriate 

```{r spline regression, echo=FALSE}
#I'm using the thin plate regression spline which is the default as it is most #popularly used spline
fit <- gam(SG ~ s(phthalates) + s(mothers_age) + s(mother_bmi) + (gest_age_weeks) + (time.cos) + (time.sin), data = SG_sub)
summary(fit)
#all of the edfs are lower than 8 so linear model is better fit than nonlinear #model


#EDF Values of around 1 tend to be close to a linear term. Anything at 8 or above is non-linear 
#as shown here: https://kevintshoemaker.github.io/NRES-746/Generalized%20Additive%20Models%20(GAMs).pdf
#date of visit and phtalates have a significant association with SG

#subset only to variables needed for imputation model
SG_sub<-SG_sub[,.(folio, SG, phthalates, mothers_age, mother_bmi, gest_age_weeks, time.sin, time.cos)]

```
## Use Mice package to run multiple imputation

```{r multiple imputation, echo=FALSE}

#MICE PACKAGE
#call the mice package
init = mice(SG_sub, maxit=0) 
meth = init$method
predM = init$predictorMatrix

#indicate the variables that will be removed as predictors
predM[, c("SG", "folio")]=0


#indicate the predictors and specify the methods for imputing the missing values
#you can create a "meth" vector that assigns a regression to each variable
#and then have "method = meth" in the multiple imputation model
#But Elena said it was okay to use default which is method=NULL
#meth[c("phthalates")]="norm"
#meth[c("mothers_age")]="norm"
#meth[c("mother_bmi")]="norm"
#meth[c("gest_age_weeks")]="norm"
#meth[c("time.cos")]="norm"
#meth[c("time.sin")]="norm"

#run the multiple (m=10) imputation model
#instead of using default, I used "cart" method which stands for 
#"classification and regression trees"

set.seed(103)
imputed = mice(SG_sub, method='cart', predictorMatrix=predM, m=10)
#CORRECTION: using complete function from mice package is not what we wanted
#instead, use default method (pmm) in mice package, 
#and you should not use the complete function to apply imputation; Instead, you #should extract all bootstrapping datasets (in our case, all 10 of the #subsamples) then find the mean of all of them and apply them to imputation of #missing data
#imputed <- complete(imputed)
#The above information is for my edification. We have moved on to using
#xgboost instead of mice package as shown here: #Nia_McRae_SG_Predict_xgboost_oct30_2018.Rmd

#Check for missing data in the imputed dataset.
#great! no missing
sapply(imputed, function(x) sum(is.na(x)))

#convert to data table
class(imputed)
imputed<-as.data.table(imputed)


#Density Plot of Original and Imputed Specific Gravity
#Good! It looks aligned. Using 'cart' method was the best decision.
plot(density(imputed$SG), col="blue", 
     xlab="Specific Gravity", #Change the x-axis label
     ylab="Density", #y-axis label
     main="Density Plot of Original and Imputed Specific Gravity")#Main title
lines(density(SG_sub$SG, na.rm=TRUE), col="red")
legend(1.025, 50, legend=c("Original SG", "Imputed SG"),
       col=c("red", "blue"), lwd=1:1, cex=0.8,
       title="SG Type", text.font=4, bg='white')


#For the sake of comparison, let's impute with mean and compare density plots
set.seed(103)
imputed_mean = mice(SG_sub, method='mean', predictorMatrix=predM, m=10)
imputed_mean <- complete(imputed_mean)
imputed_mean<-as.data.table(imputed_mean)

#Imputing with mean doesn't create as much alignment as using the default method
#it's good that we used the default method and not the population mean
plot(density(imputed_mean$SG), col="blue", 
     xlab="Specific Gravity", #Change the x-axis label
     ylab="Density", #y-axis label
     main="Density Plot of Original and Imputed Specific Gravity")#Main title
lines(density(SG_sub$SG, na.rm=TRUE), col="red")
legend(1.025, 80, legend=c("Original SG", "Imputed SG"),
       col=c("red", "blue"), lwd=1:1, cex=0.8,
       title="SG Type", text.font=4, bg='white')

#for multiple imputation, "cart" method (i.e. decision trees) produced better density plots 
#than mean and default method

#I'm writing cleaned dataset for Allan and Elena

write.csv(SG_sub, file = "SG_Impute_Allan_Elena_oct23_2018.csv", row.names=FALSE)


```
```{r mice, echo=FALSE}


#FINAL STEP (PENDING: merge imputed data to main dataset, set SG outlier  
#for folio 380 to missing)

