---
title: "Nia_McRae_SG_Predict_xgboost_oct30_2018.Rmd"
author: "Nia McRae"
date: "October 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Nia's R Markdown for Multiple Imputation

DATE: October 30th, 2018

Nia McRae and Erik de Water created this R markdown for conducting multiple imputation for 207 missing Specific Gravity data

They split up the work. Below is the code for Nia's steps. Erik's steps are in a seperate R markdown document

Erik and Nia are using github to follow each other's work and will eventually merge their R markdowns into one complete document in an effort to produce reproducible code for multiple imputation. 

AMENDMENT: We are not using the mice package anymore. I repeated preliminary data cleaning steps here as shown in "Nia_McRae_SG_Impute_oct10_2018.Rmd" but the final model will be from the xgboost package. It is used for predicting values which is a more efficient and modern way of handling missing in contrast to mice (multiple imputation) package

#First, install and load require packages

```{r variable cleaning and transformations}
#install and load require packages
#install.packages("mice")
#install.packages("lubridate")
#install.packages("lattice")
#install.packages("xgboost")

library(haven)
library(data.table)
library(ggplot2)
library(lubridate)
library(nlme)
library(mgcv)
library(gamm4)
library(xgboost)
library(Hmisc)

```

## Load data and subset to variables you need for the analysis

```{r load data, echo=FALSE}

#load data
SG_data <- read_sas("J:/PM/Element/DATABASES-Mexico/DATASETS SENT/erik_nia_sg_imp_oct10_2018.sas7bdat")

SG_data <-as.data.table(SG_data)

#subset to variables needed for model
SG_sub <- SG_data[,.(folio, gest_age_weeks, etapa, mothers_age, fecha_control, mother_bmi, SG,
mEP2_Conc_ngml, mBP_Conc_ngml, miBP_Conc_ngml, mBzP2_Conc_ngml, mECPP_Conc_ngml)]

#subset only to 2T where bulk of missing SG are
SG_sub <- SG_sub[etapa =="2T"]

#see that all phtalates and SG are numeric 
class(SG_sub$mEP2_Conc_ngml)
class(SG_sub$mBP_Conc_ngml)
class(SG_sub$miBP_Conc_ngml)
class(SG_sub$mBzP2_Conc_ngml)
class(SG_sub$mECPP_Conc_ngml)
class(SG_sub$SG)


```

## Compute Geometric Mean

```{r geometric mean, echo=FALSE}

#create natural log of phtalates of interest
SG_sub[, mEP2_Conc_ngml_l := log(mEP2_Conc_ngml)]
SG_sub[, mBP_Conc_ngml_l := log(mBP_Conc_ngml)]
SG_sub[, miBP_Conc_ngml_l := log(miBP_Conc_ngml)]
SG_sub[, mBzP2_Conc_ngml_l := log(mBzP2_Conc_ngml)]
SG_sub[, mECPP_Conc_ngml_l := log(mECPP_Conc_ngml)]

#create exponential function then create variable to represent the geometric #mean of all five phtalates
e <- exp(1) 
SG_sub[, phthalates := e^(rowMeans(SG_sub[, c("mEP2_Conc_ngml_l", "mBP_Conc_ngml_l", "miBP_Conc_ngml_l", 
                                               "mBzP2_Conc_ngml_l", "mECPP_Conc_ngml_l")]))]


```

#convert date of visit into days of the year and then converting 
#no need to convert it to sine/cosine as xgboost doesn't 
#require it in contrast to mice package model

```{r date variable, echo=FALSE}

SG_sub[, date := ymd(fecha_control)]
SG_sub[, days := yday(date) ]



########DESCRIPTIVE STATS##########
#review descriptive stats of all variables for  subjects with phtalates data
#mean: 1.016 (same as median), min=1.00, max=1.03, sd: 0.006

describe(SG_sub[!is.na(phthalates)]$SG)
describe(SG_sub[!is.na(phthalates)]$days)
describe(SG_sub[!is.na(phthalates)]$mothers_age)
describe(SG_sub[!is.na(phthalates)]$mother_bmi)
describe(SG_sub[!is.na(phthalates)]$gest_age_weeks)

sd(SG_sub[!is.na(phthalates)]$SG, na.rm=TRUE)
sd(SG_sub[!is.na(phthalates)]$days, na.rm=TRUE)
sd(SG_sub[!is.na(phthalates)]$mothers_age, na.rm=TRUE)
sd(SG_sub[!is.na(phthalates)]$mother_bmi, na.rm=TRUE)
sd(SG_sub[!is.na(phthalates)]$gest_age_weeks, na.rm=TRUE)


#review descriptive stats of all variables for  subjects without phtalates data
describe(SG_sub[is.na(phthalates)]$SG)
describe(SG_sub[is.na(phthalates)]$days)
describe(SG_sub[is.na(phthalates)]$mothers_age)
describe(SG_sub[is.na(phthalates)]$mother_bmi)
describe(SG_sub[is.na(phthalates)]$gest_age_weeks)

sd(SG_sub[is.na(phthalates)]$SG, na.rm=TRUE)
sd(SG_sub[is.na(phthalates)]$days, na.rm=TRUE)
sd(SG_sub[is.na(phthalates)]$mothers_age, na.rm=TRUE)
sd(SG_sub[is.na(phthalates)]$mother_bmi, na.rm=TRUE)
sd(SG_sub[is.na(phthalates)]$gest_age_weeks, na.rm=TRUE)

SG_sub[is.na(SG) & !is.na(phthalates), .N] #101 subjects does not have SG among complete #phthalates data 
SG_sub[!is.na(SG) & !is.na(phthalates), .N] #847 subjects has SG


#create ggplots with partial transparency
#PENDING: still need to add smoothing term
ggplot(SG_sub, aes(phthalates, SG, group = folio)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  theme_bw()
ggplot(SG_sub, aes(mothers_age, SG, group = folio)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  theme_bw()
ggplot(SG_sub, aes(mother_bmi, SG, group = folio)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  theme_bw()
ggplot(SG_sub, aes(gest_age_weeks, SG, group = folio)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  theme_bw()
ggplot(SG_sub, aes(days, SG, group = folio)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  theme_bw()


#there is no missing data for independent variables in regards to subjects with #SG
#there is no missing phtalates among subjects with SG
SG_sub[!is.na(SG) & !is.na(phthalates), .N]
SG_sub[!is.na(SG) & is.na(phthalates), .N]

#there is no missing mother's age among subjects with SG
SG_sub[!is.na(SG) & !is.na(mothers_age), .N]
SG_sub[!is.na(SG) & is.na(mothers_age), .N]

#there is no missing mother's bmi among subjects with SG
SG_sub[!is.na(SG) & !is.na(mother_bmi), .N]
SG_sub[!is.na(SG) & is.na(mother_bmi), .N]

#there is no missing gestational age in weeks among subjects with SG
SG_sub[!is.na(SG) & !is.na(gest_age_weeks), .N]
SG_sub[!is.na(SG) & is.na(gest_age_weeks), .N]

#there is no missing days (i.e. day of year for collection date) in weeks among subjects with SG
SG_sub[!is.na(SG) & !is.na(days), .N]
SG_sub[!is.na(SG) & is.na(days), .N]

```

## Conduct spline regression to see if nonlinear model is appropriate 

```{r spline regression, echo=FALSE}
#I'm using the thin plate regression spline which is the default as it is most #popularly used spline
fit <- gam(SG ~ s(phthalates) + s(mothers_age) + s(mother_bmi) + (gest_age_weeks) + (days), data = SG_sub)
summary(fit)

#all of the edfs are lower than 8 so linear model is better fit than nonlinear #model




```

```{r assign missing values, echo=FALSE}

####assign missing values among subjects with SG to test in xgboost###

#see missing across main dataset
sapply(SG_sub, function(x) sum(is.na(x)))

#create subset that is restricted to subjects with no missing
SG_nomiss<-SG_sub[!is.na(SG)]
sapply(SG_nomiss, function(x) sum(is.na(x)))

#assign 20 random missing values in SG
set.seed(1052)
SG_nomiss[sample(1:nrow(SG_nomiss), 20), "SG"] <- NA

#Confirm the presence of missings in the dataset
sapply(SG_nomiss, function(x) sum(is.na(x)))

#TIPS FOR IMPROVING DENSITY PLOTS USING GGPLOT
#add geom_smooth term for observed and unobserved/predicted values
#use partial transparency (i.e. alpha = 0.1) to allow you to see data points atop of each other 